---
title: "DDS_UNIT_10_Linear_Regression"
output: html_document
date: "2023-03-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE}
library(tidyverse)

```

# Load in the cars dataset
``` {r}
cars = read.csv(file.choose())

```
# check the assumptions( normality, linearity, Equal SD, Independence)
```{r}
cars%>% ggplot(aes(x=Weight, y = MPG)) + geom_point()+ 
  ggtitle("Weight vs MPG scatterplot")+ 
  geom_smooth(method = "lm")
```

# I question normality so lets look at mpg < 40, and compare test to see if there is a significant difference 

```{r}
cars %>% filter(MPG<40) %>% ggplot(aes(x=Weight, y = MPG)) + geom_point()+ 
  ggtitle("Weight vs MPG scatterplot")+ 
  geom_smooth(method = "lm")

cars2 = cars %>% filter(MPG<40)
```
# fit the model MPG = Bo +B1(Weight) for both the full dataset and the dataset with mpg < 40 
```{r}
fit= lm(MPG~Weight, data = cars)
summary(fit)
confint(fit)

fit2= lm(MPG~Weight, data = cars2)
summary(fit2)
confint(fit2)
```
# There does not apear to be a signifcant difference in results so we will proceed with the full data of mpg assuming normality 

# Cross validation of SLR models 1: Mpg = Bo + B1(weight) 2: Mpg= Bo+B1(weight)+B2(weight)^2

#create train and test datasets 75% train, 25% test 
```{r}
set.seed(5)
TrainObs= sample(seq(1,dim(cars)[1]),round(.75*dim(cars)[1]),replace = FALSE)

TrainCars=cars[TrainObs,]
TestCars= cars[-TrainObs,]
```
# Model 1 prediction MSPE 
```{r}
model1_fit= lm(MPG~Weight, data = TrainCars)
summary(model1_fit)
confint(model1_fit)

model1_preds = predict(model1_fit, newdata = TestCars)
as.data.frame(model1_preds)

MSPE1= mean((TestCars$MPG - model1_preds)^2)
```
# Model 2 prediction MSPE
```{r}

cars3 = cars %>% mutate(Weight2=Weight^2)

set.seed(5)
TrainObs= sample(seq(1,dim(cars3)[1]),round(.75*dim(cars3)[1]),replace = FALSE)

TestCars= cars3[-TrainObs,]
TrainCars=cars3[TrainObs,]

model2_fit =  lm(MPG~Weight+Weight2,data = TrainCars)

model2_preds = predict(model2_fit, newdata = TestCars)
as.data.frame(model2_preds)

MSPE2= mean((TestCars$MPG - model2_preds)^2)
```

#predict Mpg for a car that weights 2000 lbs with model 2 as it had lower MSPE
```{r}

model2 =  lm(MPG~Weight+Weight2,data = cars3)
summary(model2)
confint(model2)

Predictmpg= data.frame(Weight= c(2000),Weight2= c(4000000))

predict(model2,newdata = Predictmpg, interval = "confidence")
```
# part 3: assess the relationship between mpg and horsepower 

#impute missing values of Horsepower utilizing regression model to predict the values 
#create regression model using displacement and cylinder( dataset with !is.na )
```{r}
cars4= cars %>% filter(!is.na(Horsepower))

fit3= lm(Horsepower~Displacement+Cylinders, data = cars4)
summary(fit3)
```
# Find missing values and replace them with predicted Horsepower using corresponding Displacement & Cylinder values
```{r}
cars5= cars %>% filter(is.na(Horsepower))

PredictHP= data.frame(Model= c("renault 18i","amc concord dl"), Displacement = c(100,151), Cylinders = c(4,4))

predict(fit3,newdata = PredictHP, interval = "confidence")

# impute values rounding to whole number 

cars[351,"Horsepower"] = 73
cars[371,"Horsepower"] = 92
```
# assess the relationship between mpg and horsepower with fixed data
```{r}
cars%>% ggplot(aes(x= Horsepower,y=MPG))+ geom_point()+ ggtitle("Horsepower vs MPG")+ geom_smooth(method = "lm")

fit4=  lm(MPG~Horsepower,data = cars)
summary(fit4)
confint(fit4)
```
# Predict Mpg when horsepower = 250 
```{r}
PredictMPG2= data.frame(Horsepower= c(250))

predict(fit4,newdata = PredictMPG2, interval = "confidence")
```